Sex, education, and mental health: Analyzing American census data through statistical methods, 2018 to 2021

### 2024-10-24 paper/paper.qmd
Update paper.qmd to be more concise in language

This study employs a statistical approach in R to examine the relationships between gender, educational attainment, income and reported cases of mental depression. By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.

Therefore our key questions underpinning this research include: does higher education correlate with lower rates of depression? Are women disproportionately affected by depression compared to men at the same educational level? Can we develop a model that predicts the likelihood of psychological distress [@kessler_kessler_2012] given certain characteristics of an individual such as sex, education or age? To address these questions, we analyze national census data through various statistical methods, focusing on the intersection of gender, education, and mental health outcomes [@blewett_ipums_2023].

Predictive models - even simplified ones like the one we present below - are helpful for understanding complex issues like mental health. They allow researchers and policymakers spot trends and make smarter decisions. By looking at how factors like gender and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most. Plus, predictive models help direct future research by pointing out where more exploration is needed, ensuring we keep learning and improving our approach to mental health. In a time when mental well-being is more important than ever, being able to predict and tackle what affects it is key to building healthier communities.

>>>>

This study therefore employs a statistical approach in R to examine the relationships between sex, educational attainment, income and reported cases of mental depression. An understanding of these dynamics contribute to a deeper understanding of how education and sex intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.

The key questions underpinning this research are: does higher education correlate with lower rates of depression? Are women disproportionately affected by depression compared to men at the same educational level? Can we develop a model that predicts the likelihood of psychological distress [@kessler_kessler_2012] given certain characteristics of an individual such as sex, education or age? To address these questions, we analyze national census data through various statistical methods, focusing on the intersection of gender, education, and mental health outcomes [@blewett_ipums_2023].

Predictive models - even simplified ones like the one we present below - are useful for understanding complex issues like mental health. They allow researchers and policymakers to quickly spot trends and make better decisions. By looking at how factors like sex and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most. Plus, predictive models help direct future research by pointing out where more exploration is needed, ensuring we keep learning and improving our approach to mental health. Being able to predict and tackle what affects mental health is key to building healthier communities.
